---
title: "Coverage Probability"
author: "Alan T. Arnholt"
date: 'Last edited on `r format(Sys.time(), "%B %d, %Y at %X")`'
output:
  ioslides_presentation
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

## Coverage Probability

The coverage probability of a confidence interval procedure for estimating $\pi$ at a fixed value of $\pi$ is

$$C_n(\pi) = \sum_{k=0}^nI(k, \pi)\binom{n}{k}\pi^k(1 - \pi)^{n-k}$$

where $I(k, \pi)$ equals 1 if the interval contains $\pi$ when $X = k$ and equals 0 if it does not contain $\pi$.

## Coverage Probability (concept)

* The coverage probability of a confidence interval is the proportion of all possible confidence intervals for a fixed $\pi$ that contain $\pi$.

* Confidence intervals are constructed at a given confidence level $(1 - \alpha)$, which is referred to as the **nominal coverage probability** or the **nominal confidence level**.

* In an ideal setting, the nominal confidence level will equal the coverage probability; however, when assumptions used to derive a confidence interval are not satisfied, the actual coverage probability can be either **less than** or **greater than** the nominal confidence level.

## Example 8.24

Consider a random variable $X \sim Bin(n = 25, \pi)$, and define $P = \frac{X}{n}$.

* Compute the coverage probability for a 95% Wald (asymptotic) confidence interval if $\pi = 0.70$.

* The Wald (asymptotic) confidence interval for $\pi$ is given below.

$$CI_{1 - \alpha}(\pi) = \left[ p - z_{1 - \alpha/2} \sqrt{\frac{p(1 - p)}{n}}, p + z_{1 - \alpha/2} \sqrt{\frac{p(1 - p)}{n}}\right]$$


## Solution

To compute $C_{n = 25}(\pi = 0.70)$, one must consider all the possible outcomes for $X$ when $n = 25$.  The random variable $X$ can assume values $0, 1, 2, \ldots,25$, and for each value of $X$ a different value of $p$ (the sample proportion of successes) results, which one uses with the Wald (asymptotic) confidence interval to compute a 95% confidence interval.

## Code

```{r, echo = TRUE}
n <- 25            # number of Bernoulli trials
alpha <- 0.05      # alpha level
x <- 0:n           # vector containing values RV can assume
p <- x/n           # vector of possible p values
z <- qnorm(1 - alpha/2)     # critical value
ME <- z*sqrt(p*(1 - p)/n)   # margin of error
lcl <- p - ME      # lower confidence limit  
ucl <- p + ME      # upper confidence limit  
PI <- 0.70         # PI = P(Success)
BP <- dbinom(x, n, PI)      # Binomial probability
cover <- (PI >= lcl) & (PI <= ucl)  # Logical vector 
```

## Code (continued)

```{r, echo = TRUE}
RES <- cbind(x, p, lcl, ucl, BP, cover) # cover is coerced to 0/1
DT::datatable(round(RES, 4), options = list(pageLength = 5, 
                                            autoWidth = TRUE))
```

## Computing the Coverage Probability

* Recall that $C_n(\pi) = \sum_{k=0}^nI(k, \pi)\binom{n}{k}\pi^k(1 - \pi)^{n-k}$. 
* Need to programatically add all of the Binomial Probabilities (`BP`) values when the Wald interval contains $\pi$.

```{r, echo = TRUE, comment = NA}
x[cover]
```


In this problem, $$C_{n = 25}(\pi = 0.70) = P(X = 13) + \cdots + P(X = 21)$$.

## Final Code

```{r, echo = TRUE, comment = NA}
dbinom(x[cover], n, PI)
sum(dbinom(x[cover], n, PI))
binom::binom.coverage(p = 0.70, n = 25, 
                      conf.level = 0.95, method = "asymptotic")
```

$C_{n = 25}(\pi = 0.70) = `r sum(dbinom(x[cover], n, PI))`$.


## Example 8.24 (continued)

* Compute and graph the coverage probability for the Wald (asymptotic) confidence interval, using a confidence level of 95% with 2000 equally spaced values of $\pi$.

* Previously we computed the coverage probability when $\pi$ was 0.70.  In this problem we will need to compute 2000 coverage probability values and graph those against the 2000 values of $\pi$.

## R Code

```{r, echo = TRUE, comment = FALSE}
n <- 25            # number of Bernoulli trials
alpha <- 0.05      # alpha level
CL <-  1 - alpha   # Confidence level
x <- 0:n           # vector containing values RV can assume
p <- x/n           # vector of possible p values
z <- qnorm(1 - alpha/2)     # critical value
ME <- z*sqrt(p*(1 - p)/n)   # margin of error
lcl <- p - ME      # lower confidence limit  
ucl <- p + ME      # upper confidence limit  
m <- 2000
PI <- seq(1/m, 1 - 1/m, length = m)   # PI = P(Success)
P_cov <- numeric(m) # allocating storage space
for(i in 1:m){
cover <- (PI[i] >= lcl) & (PI[i] <= ucl)  # Logical vector 
P_cov[i] <- sum(dbinom(x[cover], n, PI[i]))
}
```

## Final Graph Code

```{r, echo = TRUE,  eval = FALSE}
plot(PI, P_cov, type = "l", xlab = expression(pi), 
     ylab = "Coverage Probability", ylim = c(0.0, 1.05))
lines(c(1/m, 1 - 1/m), c(CL, CL), col = "red", 
      lty = "dotted")
text(0.5, CL + 0.05, paste("Targeted Confidence Level =", CL))
```

## Final Graph

```{r, echo = FALSE, fig.height = 5, fig.width = 6, fig.align = "center"}
plot(PI, P_cov, type = "l", xlab = expression(pi), 
     ylab = "Coverage Probability", ylim = c(0.0, 1.05))
lines(c(1/m, 1 - 1/m), c(1 - alpha, 1 - alpha), col = "red", 
      lty = "dotted")
text(0.5, 1, paste("Targeted Confidence Level =", 1 - alpha))
```

## `ggplot2` code

```{r, echo = TRUE, eval = FALSE}
DF <- data.frame(PI, P_cov)
library(ggplot2)
ggplot(data = DF, aes(x = PI, y = P_cov)) + 
  geom_line() + 
  theme_bw() + 
  labs(x = expression(pi), y = "Coverage Probability") + 
  geom_hline(yintercept = 0.95, color = "red", lty = "dashed") + 
  geom_text(aes(x = 0.5, y = 1), 
            label = paste("Targeted Confidence Level = ", CL))
```

## `ggplot2` Graph

```{r, fig.align = "center", message = FALSE, warning = FALSE}
DF <- data.frame(PI, P_cov)
library(ggplot2)
ggplot(data = DF, aes(x = PI, y = P_cov)) + 
  geom_line() + 
  theme_bw() + 
  labs(x = expression(pi), y = "Coverage Probability") + 
  geom_hline(yintercept = 0.95, color = "red", lty = "dashed") + 
  geom_text(aes(x = 0.5, y = 1), 
            label = paste("Targeted Confidence Level = ", CL))
```

## Using `binom`

```{r, echo = TRUE, fig.align = "center", message = FALSE}
library(binom)
binom.plot(n = 25, method = binom.asymp, np = 2000)
```

## Which One?

```{r, fig.align = "center", warning = FALSE, message = FALSE}
## Traditional and Agresti-Coull
n <- 20 # comment
alpha <- 0.05
x <- 0:n
p <- x/n
z <- qnorm(1 - alpha/2)
ntilde <- n + z^2
pAC <- (1/ntilde)*(x + 1/2*z^2)
m <- 10000
PI <- seq(1/m, 1 - 1/m, length = m)
m.err <- z*sqrt(p*(1 - p)/n)
m.errAC <- z*sqrt(pAC*(1 - pAC)/ntilde)
lcl <- p - m.err
lclAC <- pAC - m.errAC
ucl <- p + m.err
uclAC <- pAC + m.errAC
lengthTRAD <- ucl - lcl
lengthAC <- uclAC - lclAC
pcoverTRAD <- numeric(m)
pcoverAC <- numeric(m)
ELtrad <- numeric(m)
ELac <- numeric(m)
for(i in 1:m){
  cover <- (PI[i] >= lcl) & (PI[i] <= ucl)  # vector of 0s and 1s 
  coverAC <- (PI[i] >= lclAC) & (PI[i] <= uclAC)
  prelTrad <- dbinom(x[cover], n, PI[i])
  prelAC <- dbinom(x[coverAC], n, PI[i])
  pcoverTRAD[i] <- sum(prelTrad)
  pcoverAC[i] <- sum(prelAC)
  ELtrad[i] <- sum(lengthTRAD * dbinom(x, n, PI[i]))
  ELac[i] <- sum(lengthAC * dbinom(x, n, PI[i]))
}
# plot(PI, ELtrad, type = "l")
# lines(PI, ELac)
# Exact now
LCI <- numeric(n+1)
UCI <- numeric(n+1)
for(j in 1:(n+1)){
  LCI[j] <- binom.test(x[j], n, conf.level = 1 - alpha)$conf[1]
  UCI[j] <- binom.test(x[j], n, conf.level = 1 - alpha)$conf[2]
}
lengthEXACT <- UCI - LCI
pcoverEXACT <- numeric(m)
ELexact <- numeric(m)
for(i in 1:m){
  cover <- (PI[i] >= LCI) & (PI[i] <= UCI)  # vector of 0s and 1s 
  prelEXACT <- dbinom(x[cover], n, PI[i])
  pcoverEXACT[i] <- sum(prelEXACT)
  ELexact[i] <- sum(lengthEXACT * dbinom(x, n, PI[i]))
}
# lines(PI, ELexact, type = "l")
# Wilson Score
LCIw <- numeric(n+1)
UCIw <- numeric(n+1)
for(j in 1:(n+1)){
  LCIw[j] <- prop.test(x[j], n, conf.level = 1 - alpha, correct = FALSE)$conf[1]
  UCIw[j] <- prop.test(x[j], n, conf.level = 1 - alpha, correct = FALSE)$conf[2]
}
lengthWilson <- UCIw - LCIw
pcoverWilson <- numeric(m)
ELwilson <- numeric(m)
for(i in 1:m){
  coverWilson <- (PI[i] >= LCIw) & (PI[i] <= UCIw)  # vector of 0s and 1s 
  prelWilson <- dbinom(x[coverWilson], n, PI[i])
  pcoverWilson[i] <- sum(prelWilson)
  ELwilson[i] <- sum(lengthWilson * dbinom(x, n, PI[i]))
}
# lines(PI, ELwilson, type = "l")
####  ggplot2 now
DF <- data.frame(Interval = rep(c("Wald", "Clopper-Pearson",
                                  "Agresti-Coull", "Wilson"),
                                each = m), Pi = rep(PI, 4), 
                 ProbCoverage = c(pcoverTRAD, pcoverEXACT, 
                                  pcoverAC, pcoverWilson),
                 Elength = c(ELtrad, ELexact, ELac, ELwilson))


ggplot(data = DF, aes(x = Pi, y = ProbCoverage, color = Interval)) + 
  geom_line() + facet_wrap( ~ Interval, nrow = 2) + 
  geom_hline(yintercept = 1 - alpha, linetype = "dashed") + 
  labs(x = expression(pi), y = "Probability Coverage") + 
  theme_bw() + 
  ylim(0.7, 1.0)
```



