---
title: "Day One"
author: "Alan Arnholt"
date: Jan 17, 2017
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, comment = NA, warning = FALSE, message = FALSE, fig.align = "center")
```

# Creating A Data Frame

```{r}
ws <- c(67, 64, 66, 67, 68, 61, 64, 65, 69, 71)
hi <- c(70, 69, 72, 70, 70, 62, 68, 65, 66, 77)
name <- c("Dustin", "Cescily", "Averia", "John", "Carter", "Jaehee",
          "Paige", "Sierra", "Austin", "Lee")
sex <- c("M", "F", "F", "M", "M", "F", "F", "F", "M", "M")
DF <- data.frame(Name = name, WingSpan = ws, HeightIn = hi, Sex = sex)
# Remove individual vectors
rm(ws, hi, sex)
str(DF)
DF$Name <- as.character(DF$Name)
str(DF)
```

## Creating a Scatter Plot

```{r}
library(ggplot2)
ggplot(data = DF, aes(x = HeightIn, y = WingSpan, color = Sex)) +
  geom_point() + 
  geom_smooth(method = "lm") + 
  theme_bw()
```

## Computing Descriptive Statistics

```{r, message = FALSE}
library(dplyr)
Res <- DF %>%
  group_by(Sex) %>%
  summarize(MH = mean(HeightIn), SH = sd(HeightIn), n = n())
Res
```

## Showing Tables

```{r}
knitr::kable(Res)
```

```{r}
DT::datatable(Res)
```

## Testing Hypotheses

$H_0: \textrm{The average height of males is the no different than the average height of females.}$

$H_A: \textrm{The average height of males is greater than the average height of females.}$

Written mathematically, 

$H_0: \mu_{\textrm{males}} - \mu_{\textrm{females}} = 0$

$H_A: \mu_{\textrm{males}} - \mu_{\textrm{females}} > 0$

* Question - How should be test our hypotheses?

1. $t$-test?
2. Permutation Test?
3. Bootstrap CI?

* What are the assumptions for a $t$-test?

```{r}
# Using ggplot2
ggplot(data = DF, aes(sample = HeightIn, color = Sex)) +
  geom_qq() + 
  theme_bw()
```

```{r}
xbar1 <- Res[2, 2]
xbar2 <- Res[1, 2]
s1 <- Res[2, 3]
s2 <- Res[1, 3]
n1 <- Res[2, 4]
n2 <- Res[1, 4]
TS <- (xbar1 - xbar2)/(s1^2/n1 + s2^2/n2)^0.5
names(TS) <- "TestStatistic"
TS
str(TS$TestStatistic)
nu <- (s1^2/n1 + s2^2/n2)^2 / ((s1^2/n1)^2/(n1-1) + (s2^2/n2)^2/(n2-1))
names(nu) <- "DegreesOfFreedom"
nu
```

During class, the question was asked "What are degrees of freedom?"  A good explanation is provided on [Wikipedia](https://en.wikipedia.org/wiki/Degrees_of_freedom_(statistics)).  

## Base R

```{r}
curve(dt(x, nu$DegreesOfFreedom), from = -4, to = 4, n = 500, col = "purple", ylab = "")
```

## Curve with `ggplot2`

```{r}
ggplot(data.frame(x = c(-4, 4)), aes(x=x)) +
  stat_function(fun = dt, args = list(df = nu$DegreesOfFreedom)) + 
  theme_bw() + 
  geom_vline(xintercept = TS$TestStatistic, color = "purple") + 
  geom_hline(yintercept = 0)
```

## $p$-value

```{r}
ts <- TS$TestStatistic
ts
nu <- nu$DegreesOfFreedom
nu
pvalue <- pt(ts, nu, lower = FALSE)
pvalue
```

The $P(t_{\nu} = t_{`r nu`} \ge `r ts`) = `r pvalue`.$

```{r}
t.test(HeightIn ~ Sex, data = DF, alternative = "less")
```

## Permutation/Randomization Testing

```{r}
N <- 10^4 - 1
PTS <- numeric(N)
for(i in 1:N){
  PTS[i] <- t.test(HeightIn ~ sample(Sex), data = DF)$stat
}
epv <- (sum(PTS >= ts) + 1)/(N + 1)
epv
ggplot(data = data.frame(x = PTS), aes(x = x)) +
  geom_density(fill = "purple") + 
  theme_bw()
```

## Bootstrapping Percentile CI

* Key difference is that sampling is done with replacement!

```{r}
HeightInM <- DF$HeightIn[DF$Sex =="M"]
HeightInM
HeightInF <- subset(DF, Sex == "F", select = HeightIn, drop = TRUE)
HeightInF
```

```{r}
B <- 10^4 - 1
BSD <- numeric(B)
for(i in 1:B){
  bs1 <- sample(HeightInM, size = 5, replace = TRUE)
  bs2 <- sample(HeightInF, size = 5, replace = TRUE)
  BSD[i] <- mean(bs1) - mean(bs2)
}
ggplot(data = data.frame(x = BSD), aes(x = x)) +
  geom_density(fill = "lightblue") +
  theme_bw()
quantile(BSD, probs = c(0.05, 0.90))
```